---
title: "Getting Started With Tableau"
author: "Liu Chaoran"
date: "10 August 2015"
output: html_document
---
## Intro to Tableau
Aspired by the course 'Data Visualization' offered by University of Illinois on Cousera, I have worked on the interactive data visualization using Tableau. There is a free version of [Tableau Public](https://public.tableau.com/s/) is available and you can upload the visualization online for sharing.    
Tableau is one of the Business Intelligence tools that makes it easier to do with aesthetic chart plotting and interactive report generating. There are 3 main components used in Tableau: Worksheet, Dashboard and Story.    
* __Worksheets__ are single chart or plot
* __Dashboard__ is a single page can compose with mupliple charts or plots
* __Story__ is like powerpoint in MS Office, which put a series of pages of charts in sequence.    

There are two kinds of charts that interest me: smart map and bubble chart.   
I'm going to make use of the two chart to visual the U.S. Weather data set.

## Description of the data set
The data is taken by U.S. Weather Serivce, the link of data set is available [here](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2). The comphrensive description of the data is in [here](https://d396qusza40orc.cloudfront.net/repdata%2Fpeer2_doc%2Fpd01016005curr.pdf).   

## Data manipulation using R
As Tableau is not ideal for data munging, I would like to do the first-hand pre-processing with R. 

```{r,comment=NA}
head(data[c("EVTYPE",   "LATITUDE"  , "LONGITUDE" , "STATE" ,"BGN_DATE","FATALITIES","INJURIES","PROPDMG","PROPDMGEXP" )])
```

### Formatting variable Year, Longitude, Latitude
```{r,comment=NA}
data$BGN_DATE<-as.Date(data$BGN_DATE,format='%m/%d/%Y %H:%M:%S')
library(lubridate)
data$Year<-year(data$BGN_DATE)
data$LONGITUDE<-(-data$LONGITUDE/100)
data$LATITUDE<-data$LATITUDE/100
head(data[c("EVTYPE",   "LATITUDE"  , "LONGITUDE" , "STATE" ,"Year","FATALITIES","INJURIES","PROPDMG","PROPDMGEXP" )])
```

### Creating Economic Loss = Property Loss + Crop Loss
```{r}
data$PROPDMGEXP<-transformEXP(data$PROPDMGEXP)
data$CROPDMGEXP<-transformEXP(data$CROPDMGEXP)
data$economicLoss<-with(data,(PROPDMG*(10**PROPDMGEXP)+CROPDMG*(10**CROPDMGEXP)))
```
### Creating Health Impact = Fatalities + Injuries
```{r,comment=NA}
data$healthImpact<-data$FATALITIES+data$INJURIES
head(data[c("EVTYPE",   "LATITUDE"  , "LONGITUDE" , "STATE" ,"BGN_DATE","healthImpact","economicLoss" )])
```

### Subset the top 5 weather
Too many EVTYPE, let's look at the top 5 in frequency, after removing the disasters without impact or loss.
```{r,comment=NA}
data<-data[data$economicLoss!=0 | data$healthImpact!=0,]
freqEVTYPE<-data.frame(table(data$EVTYPE))
freqEVTYPE<-sortDF(freqEVTYPE,freqEVTYPE$Freq)
head(freqEVTYPE)
top5<-as.character(freqEVTYPE[1:5,1])
data<-data[data$EVTYPE %in% top5,]
```

### Select the variables we need
```{r}
var<-c("EVTYPE",   "LATITUDE"  , "LONGITUDE" , "STATE" ,"Year","healthImpact","economicLoss" )
data<-data[,var]
```

### Output the data.csv
```{r,eval=FALSE}
write.csv(data,'data.csv',row.names=F)
```
