---
title: "Tips of R Markdown Report"
author: "Chaoran Liu"
date: "`r Sys.Date()`"
output: 
  rmdformats::readthedown:
    code_folding: show
    highlight: kate
    toc_depth: 2
    number_sections: no
    self_contained: true
---
<style>
  #content {
    max-width: 1400px;
  }
  #sidebar h2 {
    background-color: #008B8B;
  }
  
  h1, h2 {
    color: #008B8B;
  }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.width = 8,
                      fig.height = 4)
library(dplyr)
library(ggplot2)
library(plotly)
library(googleVis)
library(knitr)
library(kableExtra)
library(DT)
library(summarytools)
library(forcats)
library(ggpubr)
```

# Introduction
<hr>

When presenting the data overview and exploratory analysis results, we used to copy a lots tables, charts from Rstudio to PowerPoint, which makes the presentation preparation painful. It become essential for data scientist to make use of better reporting tools, such as R markdown, jupyter notebooks to author analysis presentation in a more efficient and organized way, of course, we also want this to be reproducible!

In this post, I would like to share some tips when I explore building analysis report using R markdown/notebook. 

# R markdown

<hr>

Yihui Xie provided a very comprehensive and updated version of R markdown cookbook at https://bookdown.org/yihui/rmarkdown/, which gives good explanation how to make table of content, fold code snippets, configure the yaml header and more.

# Tabs {.tabset}

<hr>

One thing I found specially useful is the tabbed section. Tab layout helps to condense the parallel and lengthy content in the report.

Simply put `{.tabset}` tag after the markdown header and the sub-headers will become the tabs. The following code snippet gives an example

```
# Tabs {.tabset}

## Header2 - Tab1
this is tab1

## Header2 - Tab2
this is tab2
```

## Header2 - Tab1
this is tab1

## Header2 - Tab2
this is tab2

<br>

# Tables

<hr>

The native markdown table isn't very user-friendly, so we have to make use of functions such as `knitr::kable` or `DT::datatable` to render the table from `data.frame`.

I would like to share some tips on choosing between `kable` and `datatable`. 

* `kable` has simpler syntax and give more appealing "table like" tables in most themes.
* `datatable` has more capability such as paged tables with download buttons. There are more configurations could be referred from its JavaScript API specifications.

In a nutshell, `kable` is preferable for smaller tables, while `datatable` is preferable for bigger tables.

## markdown table

some random markdown table

```
| Tables   |      Are      |  Cool |
|----------|:-------------:|------:|
| col 1 is |  left-aligned | $1600 |
| col 2 is |    centered   |   $12 |
| col 3 is | right-aligned |    $1 |

```

| Tables   |      Are      |  Cool |
|----------|:-------------:|------:|
| col 1 is |  left-aligned | $1600 |
| col 2 is |    centered   |   $12 |
| col 3 is | right-aligned |    $1 |

## kable

`kable` is from `knitr` package

```{r kable}
require(knitr)
require(kableExtra)
mtcars %>%
  head() %>%
  kable(digits = 1, caption = 'example of kable table') %>%
  kable_styling(full_width = FALSE, position = 'left') %>%
  row_spec(0,
           bold = T,
           color = 'white',
           background = 'black')
```

## datatable

`datatable` is from `DT` package

### JS - DataTables

options list: https://datatables.net/reference/option/

### R - DT package

https://rstudio.github.io/DT/

```{r}
mtcars %>%
  datatable(
    class = 'cell-border stripe',
    extensions = 'Buttons',
    options = list(
      dom = 'Bfrtip',
      buttons = c('copy', 'csv', 'print')
    )
  )
```


# Data Summary

<hr>

Data quality is often required to check before any real analytics works. Since this is very routine job, there's a package called `summarytools` well handles this. The following beautiful data summary report is generated for the example dataset `mtcars` using `summarytools`.

* `summartools` github page: https://github.com/dcomtois/summarytools

```{r results='asis', warning=FALSE}
require(summarytools)
mtcars %>%
  dfSummary(style = 'grid',
            graph.magnif = 0.75,
            plain.ascii = F,
            valid.col = FALSE,
            tmp.img.dir = "/tmp") %>%
  print()
```


# Static Plots

`ggplot2` is our best friend in R visualization and it has good support in R markdown. Chaining functions using `%>%` and `+` makes the code chunk beautiful!

A lot of times, we would like combined many sub-plots into one. `ggplot2::facet_grid` could do some of jobs, but I found `ggpubr::ggarrange` is more powerful that allow you to combined any plots and even tables. It's cool to put chart and table side by side. (example is given in subsequent section)

`ggrigdes` is another useful `ggplot` extension that plots multiple density plots in a single chart. This is often used when comparing profiles between groups. check the detail from here: https://cran.r-project.org/web/packages/ggridges/vignettes/introduction.html 

<hr>

## ggplot

```{r}
require(ggplot2)
cor(mtcars) %>%
  as.data.frame() %>%
  tibble::rownames_to_column('var1')  %>%
  tidyr::pivot_longer(-var1, names_to = 'var2', values_to = 'cor') %>%
  filter(var1 <= var2) %>%
  ggplot(aes(x = var1, y = var2, fill = cor, label = round(cor,2))) +
  geom_tile() +
  geom_text() +
  scale_fill_gradient2() +
  labs(title = 'example of ggplot2 in R markdown')
```

## ggpubr

combine mulitple charts or tables

```{r warning=FALSE}

require(ggpubr)
require(forcats)
# add .groups = 'drop' to remove some warnings from `dplyr`
data <- mtcars %>%
  group_by(gear) %>%
  summarise(n = n(), .groups = 'drop') %>%
  ungroup() %>%
  mutate(gear = fct_rev(factor(gear)))

plt <- data %>%
  ggplot(aes(x  = gear, y = n)) +
  geom_bar(stat = 'identity', fill = 'lightblue') +
  coord_flip() +
  labs(title = 'example of combined table and plot using ggpubr')

tbl <- ggtexttable(data, rows = NULL)
ggarrange(plt, tbl, ncol =2 , nrow = 1, widths = c(2,1))
```


# Interactive Plots

<hr>

This is the section that becomes tricky. Interaction plots are only supported in HTML R document and there is no dominating interactive visualization packages in R environment. 

* `plotly` provides comprehensive chart types, documentation and cross-language capability. However I personally don't like the style, syntax and toolbox at the right upper corner.

* `echarts4r` is a R interface for `Echarts` JavaScript library, which was open sourced by Baidu. I have tested the `latest version 0.3.2` and it works well with R markdown. 

* `googleVis` is a R interface for `Google Charts` JavaScript library, which was of course developed by Google. The package has a good collection of different chart types, but it has some unknown incompatibility with both R markdown and Shiny. I've found a workaround to integrate `googleVis` charts in R markdown, but it's not perfect. 

The interactive plots can't be shown in Github rendered pages, so I just paste the code in following sections. They should be working fine on offline R markdown.

## Plotly

Plotly R document site: https://plotly.com/r/

```{r}
require(plotly)
p <- mtcars %>%
  mutate(am = factor(am)) %>%
  ggplot(aes(x = mpg, y = wt, color = am)) +
  geom_point() +
  labs(title = 'example of plotly in R markdown')
ggplotly(p)
```

## Echarts

* echarts4r github: https://echarts4r.john-coene.com/
* Echarts site: https://echarts.apache.org/en/index.html

```{r}
library(echarts4r)
df <- data.frame(
  x = LETTERS[1:5],
  y = runif(5, 1, 5),
  z = runif(5, 3, 7)
)

df %>% 
  e_charts(x) %>% 
  e_radar(y, max = 7, name = "radar") %>%
  e_radar(z, max = 7, name = "chart") %>%
  e_tooltip(trigger = "item")
```

## Google Charts

* googleVis: https://github.com/mages/googleVis
* Google Charts: https://developers.google.com/chart

`self_contained: false` is required for googleVis charts render in R markdown, refer to the github issue [here](https://github.com/mages/googleVis/issues/92).

```
output: 
  html_document:
    self_contained: false
```

```{r results='asis'}
# self_contained: false is required for googleVis charts render in R markdown
# `results = 'asis'` is required in code snippet of R markdown
suppressPackageStartupMessages(library(googleVis))
op <- options(gvis.plot.tag="chart")
plot(gvisHistogram(dino))
```

